Limits on constraining galaxy profiles with current and future strong gravitational lensing

### 2013-02-14 limits/ms.tex
Wrote introduction.

@@

\begin{abstract}
We study the effectiveness of a variety of strong lensing scenarios to inform
the reconstruction of four simulated triaxial halos with stellar and dark
matter components. We conclude that 
1) lenses with multiple sources and high redshift contrast perform very well
2) lenses where the images are already probing the stellar dominated region 
   of a galaxy can be constrained by an estimate of the stellar mass and
   suggest the fraction of dark matter
3) if velocity dispersion measurements are to be used then a sufficiently large aperture 
   is necessary
4) lensing can be used to infer the velocity anisotropy.
We also introduce GLASS, an extensible framework for lens modeling and analysis.

@@

%We conclude that redshift contrast can significantly constrain the mass
%profiles.  Stellar mass estimates can also lead to better reconstructions by
%placing lower bounds on the mass distribution in the inner regions of galaxies.
%Velocity dispersions alone are unlikely to usefully constrain the profile
%because the profile is highly sensitive to the dispersion measurement.  
%Combined with other techniques, however, velocity dispersions can help to
%constrain the range of possible reconstructed solutions. We also introduce
%GLASS, an extensible framework for lens modeling and analysis.  
\end{abstract}
\section{Introduction}

2022 - LSST - ~5000 multiply-imaged galaxies, ~150 AGN or quasars.
2020 - SKA - 10,000 lenses possibly with central images

>>>>

\newcommand{\Glass}{{\sc Glass}}

@@

\begin{abstract}
We show the effectiveness of reconstructing the mass distribution of four mock galaxies using
a range of strong gravitational lensing assumptions. In particular, we study
the affects of the lens morphology (double or quad) with and without a central
image, angular image distribution, and red shift contrast between two sources.
The lens modeling is performed using a new free-form framework called \Glass.
We conclude that 
1) Time delays are unnecessary with strong red shift contrast,
2)

%We study the effectiveness of a variety of strong lensing scenarios to inform
%the reconstruction of four simulated triaxial halos with stellar and dark
%matter components. We conclude that 
%1) lenses with multiple sources and high redshift contrast perform very well
%2) lenses where the images are already probing the stellar dominated region 
%   of a galaxy can be constrained by an estimate of the stellar mass and
%   suggest the fraction of dark matter
%3) if velocity dispersion measurements are to be used then a sufficiently large aperture 
%   is necessary
%4) lensing can be used to infer the velocity anisotropy.
%We also introduce GLASS, an extensible framework for lens modeling and analysis.

@@

%We conclude that redshift contrast can significantly constrain the mass
%profiles.  Stellar mass estimates can also lead to better reconstructions by
%placing lower bounds on the mass distribution in the inner regions of galaxies.
%Velocity dispersions alone are unlikely to usefully constrain the profile
%because the profile is highly sensitive to the dispersion measurement.  
%Combined with other techniques, however, velocity dispersions can help to
%constrain the range of possible reconstructed solutions. We also introduce
%GLASS, an extensible framework for lens modeling and analysis.  
\end{abstract}

\section{Introduction}

To date only about 100 strong lenses are known, but within the next ten years
new surveys will discover thousands of gravitational lenses. For example, SKA
in 2020 estimates about 10,000 lenses and the LSST in 2022 expects to find
about 5,000 multiply-imaged galaxies and about 150 AGNs or quasar lenses.

A typical lens, where only the measured observables are the image positions,
can been used to estimate the total mass of a galaxy including parts of the
dark matter halo. The enclosed mass inside the Einstein radius is simple to
calculate and quite robust given the observed image positions. The exact mass
distribution is more difficult to determine, however, as there are typically
not enough constraints of the mass profile. The known degeneracies have been
discussed at length in the literature. At the very least, one would like to
have a measure of the enclosed mass at at least two distinct radii to be able
to distinguish between analytic mass models. This would break the steepness
degeneracy, also known as the mass sheet degeneracy.

Other than using the Einstein radius, the enclosed mass can be estimated
from stellar velocity dispersions. If the radius from this measurement
is sufficiently different from the Einstein radius, then this provides the
necessary data. Another solution, is to find a galaxy that is lensing two
sources. The images from each source would provide an enclosed mass. The
extent to which the radii differ is dependent upon the redshifts of the
two sources: the larger the difference the larger the difference in radii.

If one assumes knowledge of the Hubble constant $H_0$ and time delays have been
observed for a lens. In principle, this determines that mass, but in practice
this can only place strong constraints on the mass distribution, as it does not
solve the problem that the exact distribution is highly degenerate. One can
choose to postpone this problem and simply fit models to the given data.  The
earliest examples were singular isothermal sphere models (SIS) but have since
evolved to multi-component models include separate models for the stars and
dark matter. Other methods that do not ignore this problem use genetic
algorithms or linear programming to explore the large space of solutions.

Worse, though, is when only time delays are known and $H_0$ is to be inferred.
In this case, one is again free to change the mass and $H_0$ while keeping
all observables fixed. A single lens can say almost nothing about the actual
value of $H_0$. However, it has been shown that combining the modeling results
of several strong lenses with time delays can significantly reduce the allowed
values and provides a measure to within 10 percent.
